---
title: "The State of Guns in Valorant"
subtitle: "The reality of precise gunplay"
author: 
  - Bo Phi 
thanks: "Code and data are available at: https://github.com/bophiwork/paper2work."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "Advertised as an FPS shooter with precise gunplay, Valorant has quickly become of the most popular games in the world since its release in 2020. The various guns used within the game play an important role in determining the way players approach the game and the general rhythm of a match. The properties and statistics of each gun create distinct playstyles and scenarios in which each gun should be used and analyzing the data can greatly improve a player’s ability to adapt to the game and improve their game sense. This statistical report uses data obtained by Valorant to (1) examine the optimal gun for specific scenarios in Valorant, and (2), discuss the difference between guns and determine which guns are better for specific playstyles and game states. As a game with 15 million monthly active users, having an understanding of how to optimize your gameplay can provide a competitive advantage against other players when climbing the ranks."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(tidyverse)
library(palmerpenguins)
library(tidyverse)
library(palmerpenguins)
library(tinytex)
library(tidyverse)
library(knitr)
library(kableExtra)
library(dplyr)
library(here)
library(bookdown)
library(ggplot2)

gunData <- readr::read_csv("/cloud/project/inputs/data/gunData.csv"
                      )
```

# Introduction

You can and should cross-reference sections and sub-sections. For instance, Section \@ref(data). R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point). 



# Data

Our data is of penguins (Figure \@ref(fig:bills)).

```{r bills, fig.cap="Bills of penguins", echo = FALSE}

```

Talk more about it.

Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)

```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}


```

Talk way more about it. 



# Results

# Discussion

## Phantom vs. Vandal


## Playstyle Matters

## Skill Matters

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}


# Additional details


\newpage


# References


